---
title: "Question 3"
author: "Sarah Fazio"
date: "2023-02-01"
output: pdf_document
---

Write a function to generate data for this problem with parameter μ. There are K = 2 equal-sized clusters with one cluster sampled from N(−μ,σ2) and the other from N(μ,σ2), where σ2 = 1. Assume n1 = n2 = 3000 observations from each (Mclus will start to have problems for larger n, but Python should be able to handle somewhat larger sample sizes, and K-means can easily handle much larger n). Estimate GMM and K-means models for μ = 0.5, 1 and 2. Report the estimated means and variances. Discuss the results, especially the biases discussed in class and how they are affected by the separation of the means.


```{r}
library(factoextra)
library(cluster)
library(mclust)
library(lattice)
```


```{r}
##generating data

#give a mean, standard deviation, and number of observations as inputs
normal <- function(u, sd, n) {
  
  #create one distribution/cluster with positive mu
  dist1 <- rnorm(n, u, sd)
  
  #create one distribution/cluster with negative mu
  dist2 <- rnorm(n, -u, sd)
  
  #combine these two clusters into one data set
  c(dist1, dist2)
}

```

KMeans -- how to visualize??
```{r}
#printing outputs
km_means <- matrix(, nrow=3,ncol = 2)
km_var <- matrix(,nrow=3,ncol = 2)

gmm_means <- matrix(, nrow=3,ncol = 2)
gmm_var <- matrix(,nrow=3,ncol = 2)

gmm_var_2 <- c(0, 0, 0)


count <- 0

set.seed(8857436)

##parameters for data generation
num_observ_1 = 3000
num_observ_2 = 3000
sd_1 = 1
sd_2 = 1

input_parameter = c(0.5, 1, 2)

for(i in 1:length(input_parameter)) {
  
  
  #increment count for placing/summarizing output
  count <- count + 1
  
  #generating data
  dist1 <- rnorm(num_observ_1, input_parameter[i], sd_1)
  dist2 <- rnorm(num_observ_2, -input_parameter[i], sd_2)

  #combining the data for each mu
  data = c(dist1, dist2)
  
  #running k-means on the data
  clus_km <- kmeans(data, centers = 2)
  
  #running Gaussian mixture
  clus_gmm <- Mclust(data, G= 2)
  
  #hist(data)
  #rug(data, col = clus$cluster)
  #plot(data, clus$cluster)
  #plot(data, jitter(clus$cluster))
  
  #adding means and variance for each cluster for kmeans
  km_means[count,1] = mean(data[clus_km$cluster == 1])
  km_var[count,1] = sd(data[clus_km$cluster == 1])^2
  km_means[count,2] = mean(data[clus_km$cluster == 2])
  km_var[count,2] = sd(data[clus_km$cluster == 2])^2
  
  #reporting means and variance for each cluster for Gaussian
  gmm_means[count,1] = clus_gmm$parameter$mean[1]
  gmm_var[count,1] = sd(data[clus_gmm$classification == 1])^2
  gmm_means[count,2] = clus_gmm$parameter$mean[2]
  gmm_var[count,2] = sd(data[clus_gmm$classification == 2])^2

  gmm_var_2[count] = clus_gmm$parameter$variance$sigmasq

}


```


```{r}
## adding $parameter$variance$sigmasq
#gmm_var_2 <- c(0, 0, 0)

#gmm_var_2[count] = clus_gmm$parameter$variance$sigmasq

 # x = c(dist1, dist2)
  # y = c(rep(0,200), rep(1,200))
  # densityplot(~x|as.factor(y), layout=c(1,2), n=200, width=1.5)
  # densityplot(~x, n=200, width=1.5)

gmm_var_2
```


Mean of 0.5

```{r}
cat("For input mean of", input_parameter[1], "and cluster 1, kmeans means is: \n",
          km_means[1, 1], "  \n")

cat("For input mean of", input_parameter[1], "and cluster 2, kmeans means is: \n",
          km_means[1, 2], "  \n")

cat("For input mean of", input_parameter[1], "and cluster 1, kmeans variance is: \n",
          km_var[1, 1], "  \n")

cat("For input mean of", input_parameter[1], "and cluster 2, kmeans variance is: \n",
          km_var[1, 2], "  \n")
```

```{r}
cat("For input mean of", input_parameter[1], "and cluster 1, gmm means is: \n",
          gmm_means[1, 1], "  \n")

cat("For input mean of", input_parameter[1], "and cluster 2, gmm means is: \n",
          gmm_means[1, 2], "  \n")

cat("For input mean of", input_parameter[1], "and cluster 1,gmm variance is: \n",
          gmm_var[1, 1], "  \n")

cat("For input mean of", input_parameter[1], "and cluster 2, gmm variance is: \n",
          gmm_var[1, 2], "  \n")
```



Mean of 1.0

```{r}
cat("For input mean of", input_parameter[2], "and cluster 1, kmeans means is: \n",
          km_means[2, 1], "  \n")

cat("For input mean of", input_parameter[2], "and cluster 2, kmeans means is: \n",
          km_means[2, 2], "  \n")

cat("For input mean of", input_parameter[2], "and cluster 1, kmeans variance is: \n",
          km_var[2, 1], "  \n")

cat("For input mean of", input_parameter[2], "and cluster 2, kmeans variance is: \n",
          km_var[2, 2], "  \n")
```

```{r}
cat("For input mean of", input_parameter[2], "and cluster 1, gmm means is: \n",
          gmm_means[2, 1], "  \n")

cat("For input mean of", input_parameter[2], "and cluster 2, gmm means is: \n",
          gmm_means[2, 2], "  \n")

cat("For input mean of", input_parameter[2], "and cluster 1,gmm variance is: \n",
          gmm_var[2, 1], "  \n")

cat("For input mean of", input_parameter[2], "and cluster 2, gmm variance is: \n",
          gmm_var[2, 2], "  \n")
```




Mean of 2.0

```{r}
cat("For input mean of", input_parameter[3], "and cluster 1, kmeans means is: \n",
          km_means[3, 1], "  \n")

cat("For input mean of", input_parameter[3], "and cluster 2, kmeans means is: \n",
          km_means[3, 2], "  \n")

cat("For input mean of", input_parameter[3], "and cluster 1, kmeans variance is: \n",
          km_var[3, 1], "  \n")

cat("For input mean of", input_parameter[3], "and cluster 2, kmeans variance is: \n",
          km_var[3, 2], "  \n")
```




```{r}
cat("For input mean of", input_parameter[3], "and cluster 1, gmm means is: \n",
          gmm_means[3, 1], "  \n")

cat("For input mean of", input_parameter[3], "and cluster 2, gmm means is: \n",
          gmm_means[3, 2], "  \n")

cat("For input mean of", input_parameter[3], "and cluster 1,gmm variance is: \n",
          gmm_var[3, 1], "  \n")

cat("For input mean of", input_parameter[3], "and cluster 2, gmm variance is: \n",
          gmm_var[3, 2], "  \n")
```



```{r}
print(km_var)
print(gmm_var)

print(km_means)
print(gmm_means)
```


```{r}
dist1 <- rnorm(3000, 0.5, 1)
dist2 <- rnorm(3000, -0.5, 1)

final_dat_check = c(dist1, dist2)
  
#running k-means on the data
clus_km <- kmeans(final_dat_check, centers = 2)
  
#running Gaussian mixture
clus_gmm <- Mclust(final_dat_check, G= 2)

clus_gmm$parameter$mean[1]

clus_gmm$parameter$mean

clus_gmm$parameter$variance$sigmasq
```


```{r, eval=FALSE}
cat("For input mean of", input_parameter[i], ", kmeans variances are: \n",
          kmeansfit_4$totss, "  \n")


  #generating the data for each mu
  #data <- normal(i, 1, 3000)
```

